---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Unzip
```{r unzip, echo=TRUE}
unzip("activity.zip")
```
Read csv file
```{r read, echo=TRUE}
act <- read.csv("activity.csv")
```
Check data
```{r check, echo=TRUE}
head(act)
summary(act)
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day

```{r steps_day, echo=TRUE}
act_notNan <- subset(act, !is.na(act$steps))
act_steps_day <- setNames(aggregate(
                          x=act_notNan$steps, by=list(act_notNan$date), sum), 
                 c("Date", "Total_Steps"))

```

If you do not understand the difference between a histogram and a barplot, 
research the difference between them. Make a histogram of the total number 
of steps taken each day

```{r hist, echo=TRUE}
hist(act_steps_day$'Total_Steps', 
     col = "green", main="Steps Frequency", xlab = "Step Intervals")
```

Calculate and report the mean and median of the total number of steps taken 
per day

```{r meanMedian, echo=TRUE}
mean(act_steps_day$Total_Steps)
median(act_steps_day$Total_Steps)
```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
and the average number of steps taken, averaged across all days (y-axis)

```{r average, echo=TRUE}
act_mean <- aggregate(steps ~ interval, data = act_notNan, mean)
with(act_mean, plot(interval, steps, type="l", col="blue", main="Days Average"))

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
